---
title: "Homework 3"
author: "Cesar Dao (cad4837)"
date: "2024-02-05"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(mosaic)
library(tidyverse)
```

[Repository Link](https://github.com/cedaco/SDS315_HW3)


# Problem 1
```{r, echo=FALSE}
creatinine = read.csv("creatinine.csv")

ggplot(creatinine) + geom_point(aes(x = age, y = creatclear)) + geom_smooth(aes(x=age, y=creatclear), method='lm') + labs(x = "Age (years)", y = "Creatinine Clearance Rate (mL/min)", title = "Creatinine Clearance Rate With Respect to Age")

creat_model = lm(creatclear ~ age, data = creatinine)
print(creat_model$coefficients)
print("The linear regression equation for this data set is: y = 147.8129 - 0.6198159x")

```

## Part A
```{r, echo=FALSE}
cat(147.8129 - 0.6198159*55, "mL/minute")
```
The expected creatinine clearance rate for a 55 year old is 113.723 mL/minute. To solve this, I used the following linear equation, which represents the line of best fit on the graph above: y = 147.8129 - 0.6198159x. In this equation, Y represents the creatinine clearance rate in mL/minute and x represents age in years.

## Part B
```{r, echo=FALSE}
coef(creat_model)[2]
```

For each year, the creatinine clearance rate decreases by 0.6198159 mL/minute each year. We can see that through running the coef function for creat_model, we are given the intercept and the slope of our linear regression line. The slope that was returned was -0.6198159.

## Part C
```{r, echo=FALSE}
age_40_prediction = 147.8129 - 0.6198159*40
age_60_prediction = 147.8129 - 0.6198159*60
cat("Predicted clearance Rate at age 40:", age_40_prediction, "mL/minute\n")
cat("Predicted clearance Rate at age 60:", age_60_prediction, "mL/minute")
```

The expected creatinine clearance rate at 40 years old is 123.0203 mL/minute (y = 147.8129 - 0.6198159*40). This means that the given 40 year old with a creatinine clearance rate of 135 mL/minute is healthy for their age.

The expected creatinine clearance rate at 60 years old is 110.6239 mL/minute (y = 147.8129 - 0.6198159*60). This means that the given 60 year old with a creatinine clearance rate of 112 mL/minute is slightly less healthy than expected for their age; this given 60 year old's clearance rate is only about 1.4 mL/minute less than expected.

# Problem 2
The Beta of a stock can be best interpreted as the change in a stock's return value over a 1% change in market portfolio. Understanding Beta can help one predict the volatility of a stock, which can make and save individuals a lot of money. Beta is calculated by analyzing the relationships between stock returns and market returns The S&P 500 usually serves an indicator for market returns (independent variable), while stock returns can pretain to other stocks on the market (dependent variable). A Beta greater than 1 means that the stock being analyzed is volatile, while a Beta less than one means that the stock being analyzed is less volatile.

```{r, echo=FALSE}
stocks = read.csv("marketmodel.csv")

aapl_model = lm(AAPL ~ SPY, data = stocks)


goog_model = lm(GOOG ~ SPY, data = stocks)
mrk_model = lm(MRK ~ SPY, data = stocks)
jnj_model = lm(JNJ ~ SPY, data = stocks)
wmt_model = lm(WMT ~ SPY, data = stocks)
tgt_model = lm(TGT ~ SPY, data = stocks)


stocks_regression_table = tibble(
  Ticker = c("AAPL", "GOOG", "MRK", "JNJ", "WMT", "TGT"),
  Intercept = c(coef(aapl_model)[1], coef(goog_model)[1], coef(mrk_model)[1], coef(jnj_model)[1], coef(wmt_model)[1], coef(tgt_model)[1]),
  Slope = c(coef(aapl_model)[2], coef(goog_model)[2], coef(mrk_model)[2], coef(jnj_model)[2], coef(wmt_model)[2], coef(tgt_model)[2]),
  R2 = c(summary(aapl_model)$r.squared, summary(goog_model)$r.squared, summary(mrk_model)$r.squared, summary(jnj_model)$r.squared, summary(wmt_model)$r.squared, summary(tgt_model)$r.squared)
)

print(stocks_regression_table)
                                     

```
The above table shows the regression statistics for some popular stocks. The intercept shows the starting point for the regression line (alpha) and the slope shows the beta. R2 displays the correlation coefficient, which measures how dependent the 6 stocks show are to the SPY.

Systematic risk refers to the volatility of a stock based on market activity (SPY). The least risky stock of the one's shown is WMT (Walmart), for it has the lowest beta value, meaning that investors will lose less money if it is not a good day for the market. The most volatile stock appears to be AAPL (Apple), whose beta is slightly above 1, meaning that it is relatively volatile based on activity from the S&P 500.

```{r, echo=FALSE}
covid = read.csv("covid.csv")
```
First, we must see what the relationships look like without any sort of fitting, separated by country.
```{r, echo=FALSE}
ggplot(covid) + geom_line(aes(x = days_since_first_death, y = deaths, color = country)) + facet_wrap(.~country) + labs(title = "COVID Deaths Over Time In Italy And Spain", x = "Days Passed Since First Death", y = "Deaths")
```
To fit the data linearly, we must take the lograithm of our the death amount, which will also allow us to see our slopes and intercepts for the data in each respective country.
```{r, echo=FALSE}
model_italy = lm(log(deaths)~days_since_first_death, data = subset(covid, country == "Italy"))
model_spain = lm(log(deaths)~days_since_first_death, data = subset(covid, country == "Spain"))

cat("Intercept for Italy's line of best fit: ", round(coef(model_italy)[1], 3), "\n")
cat("Slope for Italy's line of best fit: ", round(coef(model_italy)[2], 3), "\n")
cat("Intercept for Spain's line of best fit: ",round(coef(model_spain)[1], 3), "\n")
cat("Slope for Spain's line of best fit: ",round(coef(model_spain)[2], 3))

```
With this information, we can now display graphs that have best-fit lines.

```{r, echo=FALSE}
ggplot(covid) + geom_line(aes(x = days_since_first_death, y = log(deaths), col = country)) + facet_wrap(.~country) +   
  geom_abline(data = subset(covid, country == "Italy"), aes(intercept = coef(model_italy)[1], slope = coef(model_italy)[2]), color = "black") +
  geom_abline(data = subset(covid, country == "Spain"), aes(intercept = coef(model_spain)[1], slope = coef(model_spain)[2]), color = "black") + labs(title = "Fitted Graphs For COVID Deaths Over Time In Italy And Spain", x = "Days Passed Since First Death", y = "Deaths (log(deaths))")
```
Finally we can calculate the growth rate and doubling time for each distribution. Growth rate is calculated by taking e to the power of the slope minus 1 ((e**slope) - 1). Doubling time is calculated by dividing the log(2) by the growth rate (log(2)/growth_rate). They can be seen as the following:
```{r, echo=FALSE}
italy_growth = exp(coef(model_italy)[2]) - 1
spain_growth = exp(coef(model_spain)[2]) - 1
italy_double = log(2)/italy_growth
spain_double = log(2)/spain_growth

cat("Italy Growth Rate:", round(italy_growth, 3), "\n")
cat("Italy Doubling Time:", round(italy_double), "days\n")
cat("Spain Growth Rate:", round(spain_growth, 3), "\n")
cat("Spain Doubling Time:", round(spain_double), "days")
```

# Problem 4
```{r, echo=FALSE}
milk = read.csv("milk.csv")

ggplot(milk) + geom_point(aes(x = price, y = sales))

model_milk = lm(log(sales)~log(price), data = milk)


ggplot(milk) + geom_point(aes(x = log(price), y = log(sales))) + geom_abline(aes(intercept = coef(model_milk)[1], slope = coef(model_milk)[2] ))

coef(model_milk)
cat("The elasticity of demand is: ", coef(model_milk)[2])
```
Making the graph more linear, as well as making a line of best fit allows us to see the elasticity of demand for milk. The estimated price elasticity came out to -1.618578. This shows that milk becomes less demanded when prices for it increase.